---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figure/",
  fig.height = 1
)
```

# BuenColors
A package that facilitates nice colors created for the [Buenrostro Lab](https://buenrostrolab.com). The base of the package including some divergent color schemes were taken from the [wesanderson](https://github.com/karthik/wesanderson) package. This README provides a quick overview of how to use the color schemes with ggplot2 and which palettes are available. 

## Installation

```R
devtools::install_github("caleblareau/BuenColors")
```

## Color Palettes 

```{r, library_dummy, message=FALSE, warning = FALSE}
library(BuenColors)
```

With the library loaded, just type this to get either the `continuous` or 
`discrete` (by default) palette printed in your `plot` console. 

```{r, example}
jdb_palette("FantasticFox")
jdb_palette("aqua_brick", type = "continuous")
```

### Here are all the names that are available...
```{r, palettes_dummy, echo = FALSE}
sort(names(jdb_palettes))
```

## Color Maps
To keep consistent color designations, one can use the `color_map` function to link features to their 
specific hex color annoations. For example, 

```{r, hsc1, echo = TRUE}
jdb_color_map(c("HSC"))
```

returns the hex code associated with `HSC` in the Buenrostro Lab paradigm. This function may be applied 
over multiple features--

```{r, hsc2, echo = TRUE}
jdb_color_map(c("HSC", "CMP", "HSC"))
```

and will error out when a feature is not recognized--

```{r, hsc0, echo = TRUE, eval = FALSE}
jdb_color_map(c("WHAT"))
#>  Error: all(name %in% names(jdb_color_maps)) is not TRUE 
```

### Here are all the names that are available...
```{r, colmaps_dummy, echo = FALSE}
sort(names(jdb_color_maps))
```

### Here are what the mappings look like...
```{r, colormaps, fig.height = 10, echo = FALSE}
df <- data.frame(names = names(jdb_color_maps), 
                 jdb_color_maps = unname(jdb_color_maps), stringsAsFactors = FALSE)

ggplot(df, aes(x = names, y = -1, fill = names)) +
  geom_bar(stat = "identity") + pretty_plot(fontsize = 20) + 
  scale_fill_manual(values = jdb_color_maps) + coord_flip() +
  theme(legend.position = "none")+
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = NULL, y = NULL, title = "Buenrostro Lab Cell Schemes") +
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), panel.border = element_blank())

```

## ggplot example
To coordinate a ggplot feature (e.g. data point in a scatter plot) with a particular color, [this post](https://www.biostars.org/p/204891/) was a life-saver. Specifically, we'll use a named vector
to coordinate the discrete values. Here's an example--
```{r, colormapPlotExample, fig.height = 5, echo = TRUE}
xy <- 1:7
cell <- c("GMP-A", "Ery", "CD4", "Ery", "LMPP", "ERY", "MEP")

df <- data.frame(
  xy = xy,
  cell = cell, stringsAsFactors = FALSE
)

ggplot(df, aes(x = xy, y = xy, color = cell)) +
  geom_point(size = 10) + pretty_plot() +
  scale_color_manual(values = jdb_color_maps)

```

**P.S.**-- this call will return a blank color for features that are not found in the color map ("ERY" in this example).  

### Important Note on color mappings...

The above `ggplot` command works because `jdb_color_maps` (with an 's') exists as a named vector in the `BuenColors` NAMESPACE. The provided function (`jdb_color_map`) does not have an 's' by the way. This same syntax of supplying a named vector should work for all discrete color scale functionalities in `ggplot`. 

# Discrete colors
The trick here is to use `scale_color_manual` like it is shown here--

```{r, discrete, fig.height = 3}
library(ggplot2)
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) + 
  geom_point() + pretty_plot() + 
  scale_color_manual(values = jdb_palette("brewer_spectra"))
```

# Continuous fill
The trick here is to use `scale_color_gradientn` like it is shown here--

```{r, ggplotcont, fig.height = 3}
df <- data.frame(x = rnorm(1000), y = 0)
ggplot(df, aes(x=x, y=y, colour=x)) + geom_point() + 
  scale_color_gradientn(colors = jdb_palette("flame_light")) +
  pretty_plot()
```

# Density Plot
Best way that I've found to make the density color
function represented in the points. Thanks to 
[Kamil Slowikowski](http://slowkow.com/notes/ggplot2-color-by-density/) for
figuring this out.

```{r, ggplotdensty, fig.height = 5}
dat <- data.frame(
  x = c(
    rnorm(1e4, mean = 0, sd = 0.1),
    rnorm(1e3, mean = 0, sd = 0.1)
  ),
  y = c(
    rnorm(1e4, mean = 0, sd = 0.1),
    rnorm(1e3, mean = 0.1, sd = 0.2)
  )
)
dat$density <- get_density(dat$x, dat$y)
ggplot2::ggplot(dat) + geom_point(aes(x, y, color = density)) + 
  scale_color_gradientn(colors = jdb_palette("solar_extra")) +
  pretty_plot()
```

# Shuffle Plot Order
Quick wrapper using `shuf` to change the order of plotting points
(to a random presentation) to avoid hiding effects. 

```{r, tsne, fig.height = 5}
tdf<-paste(system.file('rds',package='BuenColors'),'basicTSNE.rds',sep='/')
df <- readRDS(tdf)
ggplot(shuf(df)) + geom_point(aes(X1, X2, color = counts)) + 
  scale_color_gradientn(colors = jdb_palette("solar_extra")) +
  pretty_plot()
```

```{r, makeEverything, echo = FALSE, eval = FALSE}
# Generate this
write.table(data.frame('jdb_palette("',sort(names(jdb_palettes)), '", type = "continuous")'), file = "choices_cont.txt",col.names = FALSE, quote = FALSE, row.names = FALSE, sep = "")

write.table(data.frame('jdb_palette("',sort(names(jdb_palettes)), '")'), file = "choices_disc.txt",col.names = FALSE, quote = FALSE, row.names = FALSE, sep = "")
```

## Continuous Colors
Here's what each palette looks like on a continuous scale. 

```{r, printCont, echo = FALSE}
jdb_palette("algae_earth", type = "continuous")
jdb_palette("aqua_brick", type = "continuous")
jdb_palette("aqua_tan", type = "continuous")
jdb_palette("berry", type = "continuous")
jdb_palette("blue_cyan", type = "continuous")
jdb_palette("BottleRocket", type = "continuous")
jdb_palette("BottleRocket2", type = "continuous")
jdb_palette("brewer_blue", type = "continuous")
jdb_palette("brewer_celsius", type = "continuous")
jdb_palette("brewer_fire", type = "continuous")
jdb_palette("brewer_green", type = "continuous")
jdb_palette("brewer_heat", type = "continuous")
jdb_palette("brewer_jamaica", type = "continuous")
jdb_palette("brewer_marine", type = "continuous")
jdb_palette("brewer_orange", type = "continuous")
jdb_palette("brewer_purple", type = "continuous")
jdb_palette("brewer_red", type = "continuous")
jdb_palette("brewer_spectra", type = "continuous")
jdb_palette("brewer_violet", type = "continuous")
jdb_palette("brewer_yes", type = "continuous")
jdb_palette("calma_azules", type = "continuous")
jdb_palette("calma_bosque", type = "continuous")
jdb_palette("calma_manudo", type = "continuous")
jdb_palette("calma_marino", type = "continuous")
jdb_palette("calma_morado", type = "continuous")
jdb_palette("calma_musgos", type = "continuous")
jdb_palette("Cavalcanti", type = "continuous")
jdb_palette("Chevalier", type = "continuous")
jdb_palette("china_basics", type = "continuous")
jdb_palette("china_dragon", type = "continuous")
jdb_palette("china_novice", type = "continuous")
jdb_palette("china_ranges", type = "continuous")
jdb_palette("china_sunset", type = "continuous")
jdb_palette("china_theory", type = "continuous")
jdb_palette("china_weirdo", type = "continuous")
jdb_palette("citric", type = "continuous")
jdb_palette("citric_yellow", type = "continuous")
jdb_palette("citrus", type = "continuous")
jdb_palette("corona", type = "continuous")
jdb_palette("cyan_brick", type = "continuous")
jdb_palette("cyan_green", type = "continuous")
jdb_palette("cyan_pink", type = "continuous")
jdb_palette("cyan_purple", type = "continuous")
jdb_palette("cyan_tan", type = "continuous")
jdb_palette("cyan_violet", type = "continuous")
jdb_palette("Darjeeling", type = "continuous")
jdb_palette("Darjeeling2", type = "continuous")
jdb_palette("dark_blue", type = "continuous")
jdb_palette("dark_citrus", type = "continuous")
jdb_palette("dark_cyan", type = "continuous")
jdb_palette("dark_violet", type = "continuous")
jdb_palette("dusk_dawn", type = "continuous")
jdb_palette("FantasticFox", type = "continuous")
jdb_palette("flame_artic", type = "continuous")
jdb_palette("flame_blind", type = "continuous")
jdb_palette("flame_flame", type = "continuous")
jdb_palette("flame_light", type = "continuous")
jdb_palette("flame_macaw", type = "continuous")
jdb_palette("flame_polar", type = "continuous")
jdb_palette("flame_volts", type = "continuous")
jdb_palette("flame_watts", type = "continuous")
jdb_palette("flame_weird", type = "continuous")
jdb_palette("flame_wings", type = "continuous")
jdb_palette("forest", type = "continuous")
jdb_palette("forest_citric", type = "continuous")
jdb_palette("forest_yellow", type = "continuous")
jdb_palette("GrandBudapest", type = "continuous")
jdb_palette("GrandBudapest2", type = "continuous")
jdb_palette("horizon", type = "continuous")
jdb_palette("horizon_extra", type = "continuous")
jdb_palette("lawhoops", type = "continuous")
jdb_palette("Moonrise1", type = "continuous")
jdb_palette("Moonrise2", type = "continuous")
jdb_palette("Moonrise3", type = "continuous")
jdb_palette("ocean_aqua", type = "continuous")
jdb_palette("ocean_brick", type = "continuous")
jdb_palette("ocean_citrus", type = "continuous")
jdb_palette("ocean_earth", type = "continuous")
jdb_palette("ocean_green", type = "continuous")
jdb_palette("ocean_pink", type = "continuous")
jdb_palette("ocean_red", type = "continuous")
jdb_palette("ocean_teal", type = "continuous")
jdb_palette("purple_baby", type = "continuous")
jdb_palette("purple_pink", type = "continuous")
jdb_palette("Royal1", type = "continuous")
jdb_palette("Royal2", type = "continuous")
jdb_palette("Rushmore", type = "continuous")
jdb_palette("samba_color", type = "continuous")
jdb_palette("samba_light", type = "continuous")
jdb_palette("samba_night", type = "continuous")
jdb_palette("solar_basic", type = "continuous")
jdb_palette("solar_blues", type = "continuous")
jdb_palette("solar_extra", type = "continuous")
jdb_palette("solar_flare", type = "continuous")
jdb_palette("solar_glare", type = "continuous")
jdb_palette("solar_rojos", type = "continuous")
jdb_palette("teal_orange", type = "continuous")
jdb_palette("teal_violet", type = "continuous")
jdb_palette("white_grove", type = "continuous")
jdb_palette("white_jungle", type = "continuous")
jdb_palette("white_mango", type = "continuous")
jdb_palette("white_orange", type = "continuous")
jdb_palette("white_tango", type = "continuous")
jdb_palette("wolfgang_basic", type = "continuous")
jdb_palette("wolfgang_extra", type = "continuous")
jdb_palette("Zissou", type = "continuous")
```

## Discrete colors
Here are the discrete color units that go into each scale. If n is small and discrete, note 
that the colors that are selected are read from left to right. 

```{r, printDiscrete, echo = FALSE}
jdb_palette("algae_earth")
jdb_palette("aqua_brick")
jdb_palette("aqua_tan")
jdb_palette("berry")
jdb_palette("blue_cyan")
jdb_palette("BottleRocket")
jdb_palette("BottleRocket2")
jdb_palette("brewer_blue")
jdb_palette("brewer_celsius")
jdb_palette("brewer_fire")
jdb_palette("brewer_green")
jdb_palette("brewer_heat")
jdb_palette("brewer_jamaica")
jdb_palette("brewer_marine")
jdb_palette("brewer_orange")
jdb_palette("brewer_purple")
jdb_palette("brewer_red")
jdb_palette("brewer_spectra")
jdb_palette("brewer_violet")
jdb_palette("brewer_yes")
jdb_palette("calma_azules")
jdb_palette("calma_bosque")
jdb_palette("calma_manudo")
jdb_palette("calma_marino")
jdb_palette("calma_morado")
jdb_palette("calma_musgos")
jdb_palette("Cavalcanti")
jdb_palette("Chevalier")
jdb_palette("china_basics")
jdb_palette("china_dragon")
jdb_palette("china_novice")
jdb_palette("china_ranges")
jdb_palette("china_sunset")
jdb_palette("china_theory")
jdb_palette("china_weirdo")
jdb_palette("citric")
jdb_palette("citric_yellow")
jdb_palette("citrus")
jdb_palette("corona")
jdb_palette("cyan_brick")
jdb_palette("cyan_green")
jdb_palette("cyan_pink")
jdb_palette("cyan_purple")
jdb_palette("cyan_tan")
jdb_palette("cyan_violet")
jdb_palette("Darjeeling")
jdb_palette("Darjeeling2")
jdb_palette("dark_blue")
jdb_palette("dark_citrus")
jdb_palette("dark_cyan")
jdb_palette("dark_violet")
jdb_palette("dusk_dawn")
jdb_palette("FantasticFox")
jdb_palette("flame_artic")
jdb_palette("flame_blind")
jdb_palette("flame_flame")
jdb_palette("flame_light")
jdb_palette("flame_macaw")
jdb_palette("flame_polar")
jdb_palette("flame_volts")
jdb_palette("flame_watts")
jdb_palette("flame_weird")
jdb_palette("flame_wings")
jdb_palette("forest")
jdb_palette("forest_citric")
jdb_palette("forest_yellow")
jdb_palette("GrandBudapest")
jdb_palette("GrandBudapest2")
jdb_palette("horizon")
jdb_palette("horizon_extra")
jdb_palette("lawhoops")
jdb_palette("Moonrise1")
jdb_palette("Moonrise2")
jdb_palette("Moonrise3")
jdb_palette("ocean_aqua")
jdb_palette("ocean_brick")
jdb_palette("ocean_citrus")
jdb_palette("ocean_earth")
jdb_palette("ocean_green")
jdb_palette("ocean_pink")
jdb_palette("ocean_red")
jdb_palette("ocean_teal")
jdb_palette("purple_baby")
jdb_palette("purple_pink")
jdb_palette("Royal1")
jdb_palette("Royal2")
jdb_palette("Rushmore")
jdb_palette("samba_color")
jdb_palette("samba_light")
jdb_palette("samba_night")
jdb_palette("solar_basic")
jdb_palette("solar_blues")
jdb_palette("solar_extra")
jdb_palette("solar_flare")
jdb_palette("solar_glare")
jdb_palette("solar_rojos")
jdb_palette("teal_orange")
jdb_palette("teal_violet")
jdb_palette("white_grove")
jdb_palette("white_jungle")
jdb_palette("white_mango")
jdb_palette("white_orange")
jdb_palette("white_tango")
jdb_palette("wolfgang_basic")
jdb_palette("wolfgang_extra")
jdb_palette("Zissou")
```
