---
title: Using BuenColors
author: Caleb Lareau
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    theme: readable
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Using BuenColors}
  %\usepackage[UTF-8]{inputenc}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figure/",
  fig.height = 1
)
```

# BuenColors
A package that facilitates nice colors created for the [Buenrostro Lab](https://buenrostrolab.com). The base of the package including some divergent color schemes were taken from the [wesanderson](https://github.com/karthik/wesanderson) package. This README provides a quick overview of how to use the color schemes with ggplot2 and which palettes are available. 

## Installation via GitHub

```R
devtools::install_github("caleblareau/BuenColors")
```

## Color Palettes 

```{r, library_dummy, message=FALSE, warning = FALSE}
library(BuenColors)
```

With the library loaded, just type this to get either the `continuous` or 
`discrete` (by default) palette printed in your `plot` console. 

```{r, example}
jdb_palette("FantasticFox")
jdb_palette("aqua_brick", type = "continuous")
```

### Here are all the names that are available...
```{r, palettes_dummy, echo = FALSE}
sort(names(jdb_palettes))
```

## Color Maps
To keep consistent color designations, one can use the `color_map` function to link features to their 
specific hex color annoations. For example, 

```{r, hsc1, echo = TRUE}
jdb_color_map(c("HSC"))
```

returns the hex code associated with `HSC` in the Buenrostro Lab paradigm. This function may be applied 
over multiple features--

```{r, hsc2, echo = TRUE}
jdb_color_map(c("HSC", "CMP", "HSC"))
```

and will error out when a feature is not recognized--

```{r, hsc0, echo = TRUE, eval = FALSE}
jdb_color_map(c("WHAT"))
#>  Error: all(name %in% names(jdb_color_maps)) is not TRUE 
```

### Here are all the names that are available...
```{r, colmaps_dummy, echo = FALSE}
sort(names(jdb_color_maps))
```

### Here are what the mappings look like...
```{r, colormaps, fig.height = 10, echo = FALSE}
df <- data.frame(names = names(jdb_color_maps), 
                 jdb_color_maps = unname(jdb_color_maps), stringsAsFactors = FALSE)

ggplot(df, aes(x = names, y = -1, fill = names)) +
  geom_bar(stat = "identity") + pretty_plot(fontsize = 20) + 
  scale_fill_manual(values = jdb_color_maps) + coord_flip() +
  theme(legend.position = "none")+
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = NULL, y = NULL, title = "Buenrostro Lab Cell Schemes") +
  theme(axis.title.x=element_blank(), axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), panel.border = element_blank())

```

## ggplot example
To coordinate a ggplot feature (e.g. data point in a scatter plot) with a particular color, [this post](https://www.biostars.org/p/204891/) was a life-saver. Specifically, we'll use a named vector
to coordinate the discrete values. Here's an example--
```{r, colormapPlotExample, fig.height = 5, echo = TRUE}
xy <- 1:7
cell <- c("GMP-A", "Ery", "CD4", "Ery", "LMPP", "ERY", "MEP")

df <- data.frame(
  xy = xy,
  cell = cell, stringsAsFactors = FALSE
)

ggplot(df, aes(x = xy, y = xy, color = cell)) +
  geom_point(size = 10) + pretty_plot() +
  scale_color_manual(values = jdb_color_maps)

```

**P.S.**-- this call will return a blank color for features that are not found in the color map ("ERY" in this example).  

### Important Note on color mappings...

The above `ggplot` command works because `jdb_color_maps` (with an 's') exists as a named vector in the `BuenColors` NAMESPACE. The provided function (`jdb_color_map`) does not have an 's' by the way. This same syntax of supplying a named vector should work for all discrete color scale functionalities in `ggplot`. 

# Discrete colors
The trick here is to use `scale_color_manual` like it is shown here--

```{r, discrete, fig.height = 3}
library(ggplot2)
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) + 
  geom_point() + pretty_plot() + 
  scale_color_manual(values = jdb_palette("brewer_spectra"))
```

# Continuous fill
The trick here is to use `scale_color_gradientn` like it is shown here--

```{r, ggplotcont, fig.height = 3}
df <- data.frame(x = rnorm(1000), y = 0)
ggplot(df, aes(x=x, y=y, colour=x)) + geom_point() + 
  scale_color_gradientn(colors = jdb_palette("flame_light")) +
  pretty_plot()
```

# Density Plot
Best way that I've found to make the density color
function represented in the points. Thanks to 
[Kamil Slowikowski](http://slowkow.com/notes/ggplot2-color-by-density/) for
figuring this out.

```{r, ggplotdensty, fig.height = 5}
dat <- data.frame(
  x = c(
    rnorm(1e4, mean = 0, sd = 0.1),
    rnorm(1e3, mean = 0, sd = 0.1)
  ),
  y = c(
    rnorm(1e4, mean = 0, sd = 0.1),
    rnorm(1e3, mean = 0.1, sd = 0.2)
  )
)
dat$density <- get_density(dat$x, dat$y)
ggplot2::ggplot(dat) + geom_point(aes(x, y, color = density)) + 
  scale_color_gradientn(colors = jdb_palette("solar_extra")) +
  pretty_plot()
```

# Shuffle Plot Order
Quick wrapper using `shuf` to change the order of plotting points
(to a random presentation) to avoid hiding effects. 

```{r, tsne, fig.height = 5}
tdf<-paste(system.file('rds',package='BuenColors'),'basicTSNE.rds',sep='/')
df <- readRDS(tdf)
ggplot(shuf(df)) + geom_point(aes(X1, X2, color = counts)) + 
  scale_color_gradientn(colors = jdb_palette("solar_extra")) +
  pretty_plot()
```


## Continuous Colors
Here's what each palette looks like on a continuous scale. 

```{r, printCont, echo = FALSE, eval = TRUE}
all <- sort(names(jdb_palettes))
names(all) <- sort(names(jdb_palettes))
lapply(all, function(p) jdb_palette(p, type = "continuous"))
```

## Discrete colors
Here are the discrete color units that go into each scale. If n is small and discrete, note 
that the colors that are selected are read from left to right. 

```{r, printDiscrete, echo = FALSE}
lapply(all, function(p) jdb_palette(p, type = "continuous"))
```
